---
title: "CO and NO2 emissions prediction"
author: "Dhiraj Kumar Reddy; Shashaank; Pallavi; Veera Sampth Kumar Reddy"
date: "2023-04-21"
output: html_document
---

### This R markdown is gonna predict the CO and NOX emissions of turbine using various regression methods and compare them to select the best model

#### First lets import the dataset and the required libraries
```{r}
data<-read.csv("data.csv",header = T)
data=data[,2:12]
#data<-read.csv("gt_2011.csv",header = T)
#data=data[,1:11]
head(data)
```

#### Lets do some EDA on the data

```{r}
ncol(data) #check the number of columns(variables) in the dataset
nrow(data) #check the number of row(variables) in the dataset
names(data) #check the name of variables
print("The statistical summary of the features are:")
summary(data) #+statistics of each variable
```

#### lets see a scatter plot of features to see any relation btw features and targets

```{r fig.height=8, fig.width=8}
pairs(data, pch = 21)
```

#### Lets look into correlation of each feature wrt one another
```{r fig.height=10, fig.width=10}
library(corrplot) #load the package to current environment
Corr <- cor(data) #calculate the correlation coefficient matrix of variables
Corr
corrplot(Corr,method = 'circle') # circle, square, ellipse, pie
```

#### Lets adjust the outliers in the data using box plot before replacing with median
```{r}
boxplot(data$AT,xlab="AT")
boxplot(data$AP,xlab="AP")
boxplot(data$AH,xlab="AH")
boxplot(data$AFDP,xlab="AFDP")
boxplot(data$GTEP,xlab="GTEP")
boxplot(data$TIT,xlab="TIT")
boxplot(data$TAT,xlab="TAT")
boxplot(data$TEY,xlab="TEY")
boxplot(data$CDP,xlab="CDP")
boxplot(data$CO,xlab="CO")
boxplot(data$NOX,xlab="NOX")
```




#### Lets seperate the data into training and testing
```{r}
y1 <- data$CO
y2=data$NOX
X_CO=data.frame(data[,1:9],y1)
N = nrow(X_CO)
testid = sample(c(1:N),round(0.2*N),replace = FALSE)
train_CO = X_CO[-testid,]
test_CO = X_CO[testid,]
nrow(train_CO)
nrow(test_CO)

X_NOX=data.frame(data[,1:9],y2)
N = nrow(X_NOX)
testid = sample(c(1:N),round(0.2*N),replace = FALSE)
train_NOX = X_NOX[-testid,]
test_NOX = X_NOX[testid,]
nrow(train_NOX)
nrow(test_NOX)
```

#### LEts build a Random Forest model for CO emissions first
```{r}
library(randomForest)
rf.CO =randomForest(y1~.,data=train_CO,importance =TRUE,na.action = na.exclude)
rf.CO
ImpData <- as.data.frame(importance(rf.CO))
ImpData
varImpPlot(rf.CO)
```

#### Lets see the predicted values of CO emissions using Test data
```{r}
#lets create a fnc to calculate R2 score
R2_score <- function(actual,predicted) {
  R2 <- 1 - (sum((actual-predicted)^2)/sum((actual-mean(actual))^2))
  return(R2)
}


y_pred_CO = predict(rf.CO, newdata = test_CO)
library(Metrics)
sprintf("the rmse of CO emissions using thethe test data is: %.3f",rmse(test_CO$y1, y_pred_CO))
sprintf("The R2 Score of CO emissions using the test data is: %.3f",R2_score(test_CO$y1, y_pred_CO))
plot(test_CO$y1, y_pred_CO)
```


#### Lets build a Random forest for NOX emissions prediction
```{r}
library(randomForest)
rf.NOX =randomForest(y2~.,data=train_NOX,importance =TRUE,na.action = na.exclude)
rf.NOX
ImpData <- as.data.frame(importance(rf.NOX))
ImpData
varImpPlot(rf.NOX)
```

#### Lets see the predicted values of NOX emissions using Test data
```{r}
#lets create a fnc to calculate R2 score
y_pred_NOX = predict(rf.NOX, newdata = test_NOX)
sprintf("the rmse of NOX emissions using the test data is: %.3f",rmse(test_NOX$y2, y_pred_NOX))
sprintf("The R2 Score of NOX emissions using the test data is: %.3f",R2_score(test_NOX$y2, y_pred_NOX))
plot(test_NOX$y2, y_pred_NOX)
```

### LEts build a Support vector Machine Regressor model on CO emissions
```{r}
## Fit SVR model and visualize using scatter plot

#Install Package
#install.packages("e1071")

#Load Library
library(e1071)
 

#Regression with SVM
svm_CO= svm(y1~.,data=train_CO)

#Predict using SVM regression
predYsvm_CO = predict(svm_CO, test_CO)

sprintf("the rmse of CO emissions using SVM on the test data is: %.3f",rmse(test_CO$y1, predYsvm_CO))
sprintf("The R2 Score of CO emissions using SVM on the test data is: %.3f",R2_score(test_CO$y1, predYsvm_CO))
plot(test_CO$y1, predYsvm_CO)
```

### LEts build a Support vector Machine Regressor model on NOX emissions
```{r}
## Fit SVR model and visualize using scatter plot

#Install Package
#install.packages("e1071")

#Load Library
library(e1071)
 

#Regression with SVM
svm_NOX= svm(y2~.,data=train_NOX)

#Predict using SVM regression
predYsvm_NOX = predict(svm_NOX, test_NOX)

sprintf("the rmse of NOX emissions using SVM on the test data is: %.3f",rmse(test_NOX$y2, predYsvm_NOX))
sprintf("The R2 Score of NOX emissions using SVM on the test data is: %.3f",R2_score(test_NOX$y2, predYsvm_NOX))
plot(test_NOX$y2, predYsvm_NOX)
```